---
title: "Pairs plot extended"
author: "Pablo Gregori"
date: "2022-10-24"
categories: [analysis, R]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


I'd like to make my own `pairs` plot, because all the versions I have found are good, but they miss a point about mosaic plots that I like. I show an example using the `mtcars` dataset, summarized as:

```{r}
summary(mtcars)
```

First of all, I change to, "factor" variables, the false numeric variables `vs`, `am`, and the convenient numeric variables `cyl`, `gear` and `carb`, and show the summary again:

```{r}
for(i in c('vs', 'am', 'cyl', 'gear', 'carb')) {
  mtcars[[i]] = factor(mtcars[[i]])
}
summary(mtcars)
```


Now, let's see how the plain `plot` function deals with it:

```{r}
plot(mtcars)
```

Scatterplots for all the variables. Not the best choice.

Now, let's see how the `pairs` function deals with it:

```{r}
pairs(mtcars)
```

Same output.

Hadley Wickham did a good job for generalizing this kind of plots in [their paper](https://vita.had.co.nz/papers/gpp.pdf).

They talk about `gpairs` function, in package `gpairs`.

```{r}
require(gpairs)
gpairs(mtcars)
```

Beautiful.

Only one objection. The mosaic plots show the marginal distributions of the $y$-axis variables, and the conditionals of the $x$-axis variable, given the values of the $y$-axis variables. Why?
